PUBPOL 543
Deliverable 1: Univariate plot

Note: The dataset used for this assignment was sourced from the US Government's open data repository: https://catalog.data.gov/

Reading-in the data from the Github repository location:
```{r}
location='https://github.com/elenapopic543/Univariate/raw/main/'
file='Data_Breach_Notifications_Affecting_Washington_Residents.csv'
link=paste0(location,file)
data1 <- read.csv(file=url(link), header=TRUE, stringsAsFactors=FALSE)
```

Examining the variables in the dataframe:
```{r}
names(data1)
```
Examining data types:
```{r}
str(data1,width = 70,strict.width='cut')
```
Examining raw values:
```{r}
head(data1$IndustryType,20) 
```
Creating a frequency table:
```{r}
absoluteT=table(data1$IndustryType,
                exclude = 'nothing')
absoluteT
```
Transforming the counts into proportions:
```{r}
prop.table(absoluteT)
```
Translating proportions into percents:
```{r}
propT=prop.table(absoluteT)*100
propT
```
Turning the frequency table into a dataframe:
```{r}
(tableFreq=as.data.frame(absoluteT))
```
Renaming columns and adding percents:
```{r}
names(tableFreq)=c("Industry","Count")
# adding percents:
tableFreq$Percent=as.vector(propT)

tableFreq
```

```{r}
library(ggplot2)
#Creating the base object:
base = ggplot(data = tableFreq, 
             aes(x = Industry,
                 y = Count))
#Adding layers to create default barplot:
plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') 
plot1
```
Adding titles:
```{r}
titleText='Data Breach Incidents by Industry in WA, 2016-2022'
sourceText='Source: DATA.GOV'

plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText) 
plot2
```

Using percent values instead of counts on Y-axis:
```{r}
base= ggplot(data = tableFreq, 
             aes(x = Industry,
                 y = Percent)) 

plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') 

plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
#Customizing the axes:
library(scales) 
plot3 = plot2 + scale_y_continuous(breaks=c(0,10,20,30,40,50),
                                   limits = c(0,50), 
                                   labels=unit_format(suffix = '%')) 
plot3

```
Adding the actual percent values (rounded to two decimal places):
```{r}
paste0(round(tableFreq$Percent,2), '%')
```
```{r}
LABELS=paste0(round(tableFreq$Percent,2), '%')

plot4 = plot3 + geom_text(vjust=0, 
                          size = 5,
                          aes(y = Percent,
                          label = LABELS))
plot4
```
Ordering the bars by height:
```{r}
tableFreq=tableFreq[order(tableFreq$Percent),]
tableFreq
```
Saving the order:
```{r}
(industriesOrd=tableFreq[order(tableFreq$Percent),'Industry'])
```
Finalizing the barplot:
```{r}
#Generating values for column labels
LABELS = paste0(round(tableFreq$Percent,2), '%')
#Specifying the base of the plot
base = ggplot(data = tableFreq, 
             aes(x = Industry,y = Percent)) 
#Ordering the bars by height:
base = base + scale_x_discrete(limits=industriesOrd) 
#Changing the theme to classic
base = base + theme_classic()
#Adding layers to the base 
plot1 = base + geom_bar(fill ="blue1",
                        stat = 'identity') 
#Adding title and caption
plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
#Customizing and labeling the values on the Y-axis with a % suffix:
plot3 = plot2 + scale_y_continuous (breaks=c(0,10,20,30,40,50),
                                   limits = c(0,50),
                                   labels=unit_format(suffix = '%')) 
#Moving the plot title to the center, and caption to left: 
plot4 = plot3 + theme(plot.caption = element_text(hjust = 0),
                      plot.title = element_text(hjust = 0.5))

#Adding column labels
plot5 = plot4 + geom_text(vjust=0, 
                          size = 3.5,
                          aes(y = Percent,
                          label = LABELS))
plot5
```

###
Deliverable 2: Bivariate Plot
In the following, We will explore the number of Washingtonians affected in data breaches by the type of the industry where data breach incidents occurred (i.e., perform a Categorical-Numerical bivariate relationship analysis)

Summarizing the variable reflecting the number of people affected by the data breach:
```{r}
summary(data1$WashingtoniansAffected)
```
Excluding the missing values:
```{r}
#data1_nona=data1[complete.cases(data1),]
```

Getting the median of exposed residents for each industry type:
```{r}
aggregate(data=data1, WashingtoniansAffected~IndustryType,median)
```
Getting the mean of exposed residents for each industry type:
```{r}
aggregate(data=data1, WashingtoniansAffected~IndustryType,mean)
```
Checking other statistics:
```{r}
# using "summary" function
tapply(data1$WashingtoniansAffected,
       data1$IndustryType, summary)
```
The above summary statistics point out that half of the data breach incidents in each industry type are affecting from roughly 1400 to 2400 WA residents. The incidents with a large number of exposed people are less common in each industry type. We will reflect these findings in the following boxplot:

```{r}
library(ggpubr)

base1=ggplot(data=data1[data1$WashingtoniansAffected>=1500,],
            aes(x=IndustryType,y=WashingtoniansAffected)) 
box1=base1 + geom_boxplot() + labs(title = "1.5k residents and above")

base2=ggplot(data=data1[data1$WashingtoniansAffected>=20000,],
            aes(x=IndustryType,y=WashingtoniansAffected))
box2=base2 + geom_boxplot() + labs(title = "20k residents and above")

base3=ggplot(data=data1[data1$WashingtoniansAffected>=100000,],
            aes(x=IndustryType,y=WashingtoniansAffected)) 
box3=base3 + geom_boxplot() + labs(title = "100k residents and above")

ggarrange(box1,box2,box3,ncol = 1)
```

Moving forward, we will focus our analysis on data breach incidents that affected 100k people or more.
```{r}
titleText2="Data breach incidents affecting over 100k residents (WA, 2016-2022)"
base4=ggplot(data=data1[data1$WashingtoniansAffected>=100000,],
            aes(x=IndustryType,y=WashingtoniansAffected)) 
box4=base4 + geom_boxplot() +
    labs(title = titleText2)
box4 + coord_flip()
```
Reordering by the median value of incidents:
```{r}
base5=ggplot(data=data1[data1$WashingtoniansAffected>=100000,],
            aes(x=reorder(IndustryType,
                          WashingtoniansAffected,
                          median),
                y=WashingtoniansAffected)) 
box5=base5 + geom_boxplot() + labs(title =titleText2) + scale_y_continuous (breaks=c(100000,500000,1000000,2000000,3000000),
limits = c(0,4000000), 
labels = c("100000","500000","1000000","2000000","3000000")) 

box5  + coord_flip()
```
Creating a histogram:
```{r}
baseHY=ggplot(data=data1[data1$WashingtoniansAffected>=100000,],
            aes(x=WashingtoniansAffected)) 
histHY=baseHY + geom_histogram(aes(fill=IndustryType), 
                              color='black') + 
  scale_x_continuous(breaks=c(100000,500000,1000000,2000000,3000000),
limits = c(0,4000000), 
labels = c("100000","500000","1000000","2000000","3000000")) + labs(title =titleText2)

histHY
```
```{r}
#Adding facets and further customizing the axes of the plot:
baseHY=ggplot(data=data1[data1$WashingtoniansAffected>=100000,],
            aes(x=WashingtoniansAffected)) 
histHY=baseHY + geom_histogram(aes(fill=IndustryType), 
                              color='black') + labs(title =titleText2) +
scale_x_continuous(breaks=c(100000,500000,1000000,2000000,3000000),
limits = c(0,4000000), 
labels = c("100000","500000","1000000","2000000","3000000")) 

histHY + facet_grid(~IndustryType)
```
Alternative:
```{r}
histHY + facet_grid(IndustryType~.) + guides(fill="none")
```
Creating a mean-error plot:
```{r}
data1[,c('IndustryType', 'WashingtoniansAffected')] %>%head(20)
```
```{r}
library(lattice)
library(plyr)
```

```{r}
Rmisc::group.CI(data=data1,
          WashingtoniansAffected~IndustryType)
```
```{r}
baseMEANs=ggplot(data1, aes(x=IndustryType,
                             y=WashingtoniansAffected)) +
        theme_classic()
pointMEANS=baseMEANs + stat_summary(fun = mean, 
                                    geom = "point")
pointMEANS 
```
Adding an error-bar:
```{r}
pointErrors=pointMEANS + stat_summary(fun.data = mean_ci,
                                      geom = "errorbar") 
pointErrors
```
Including the data in a plot:
```{r}
BarJit=pointErrors + geom_jitter(colour="darkseagreen4",
                             alpha=0.2 #transparency
                             )
BarJit
```














